---
title: "Capitalism VS Bolshevism: A Text Analysis"
author:
  - name: Conner Jainese 
    url: {}
date: 2022-03-10
output: 
  distill::distill_article:
    self_contained: false
---

### Overview

In this report we will perform a text analysis on the book: "Capitalism VS Bolshevism", published in 1919 by American author George L. Walker^1^. More specifically, we will identify the most commonly used words in each chapter, and perform a sentiment analysis of the text using the NRC Word-Emotion Association Lexicon^2^. This lexicon acts as a reference assigning one of eight basic emotions (anger, fear, anticipation, trust, surprise, sadness, joy, disgust) and two general sentiments (positive and negative) to a list of commonly used English words. The NRC Lexicon was developed via crowd sourcing.   

```{r setup, include=TRUE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# load packages
library(tidyverse)
library(here)
library(tidytext)
library(pdftools)
library(stringr)
```

```{r}
# read in the text
text <- pdf_text(here("data", "capitalismvsbolshevism.pdf"))

# make text into dataframe
text_lines <- data.frame(text) %>%
  # add column for page number and fill in
  mutate(page = 1:n()) %>% 
  # separate rows of text
  mutate(text_full = str_split(text, pattern = "\\n")) %>%
  # split up nested rows of texted into their own row of the df
  unnest(text_full) %>%
  # remove whitespaces
  mutate(text_full = str_trim(text_full))

# wrangle chapters
text_chapters <- text_lines %>% 
  # remove beginning part of book before main text
  slice(-(1:211)) %>% 
  # create column with chapter using str_detect
  mutate(chapter = ifelse(str_detect(text_full, "Chapter"), 
                          text_full, 
                          NA)) %>%
  # auto fill downwards to populate all lines in chapter
  fill(chapter, .direction = "down") %>%
  # separate word chapter from chapter number 
  separate(col = chapter, 
           into = c("ch", "no"), 
           sep = " ") %>%
  # make reglar chapter numbers from roman numerals 
  mutate(chapter = as.numeric(as.roman(no)))

# wrangle for word counts
text_words <- text_chapters %>%
  # unnest to get indiviual words on separate rows 
  unnest_tokens(word, text_full) %>% 
  # remove text column
  select(-text)

# make clean, no stop words
text_words_clean <- text_words %>%
  # remove stop words 
  anti_join(stop_words, by = "word")

# count them up
word_counts <- text_words_clean %>%
  # remove na values
  na.omit() %>% 
  # count by chapter
  count(chapter, word)

# find top 5 words
top_5 <- word_counts %>%
  # group by chapter
  group_by(chapter) %>% 
  # arrange in descending order
  arrange(-n) %>% 
  # take the top 5
  slice(1:5) %>%
  # ungroup
  ungroup()
```

```{r}
# chapter names list for plots
chapters <- c("1" = "The System", 
               "2" = "Capital", 
               "3" = "Overproduction", 
               "4" = "War Lessons", 
               "5" = "A Protective Tariff", 
               "6" = "Wages", 
               "7" = "Bolshevism", 
               "8" = "Discontent", 
               "9" = "Profits and Interest", 
               "10" = "The Corporation", 
               "11" = "Wall Street", 
               "12" = "Money", 
               "13" = "Taxation", 
               "14" = "Luxuries", 
               "15" = "Our Opportunity")

# plotting top 5
ggplot(data = top_5, 
       aes(x = n, 
           y = word)) +
  # pick fill color
  geom_col(fill = "aquamarine4") +
  # facet wrap by chapter and free scales
  facet_wrap(~chapter, scales = "free", labeller = labeller(chapter = chapters)) +
  # change theme
  theme_test() +
  # update labs
  labs(y = " ",
       x = " ", 
       title = "Top 5 Words Used Per Chapter") +
  # reduce breaks to declutter
  scale_x_continuous(n.breaks = 4) +
  # adjust title and facet titles 
  theme(plot.title = element_text(size = 18, hjust = 0.5), 
        strip.text.x = element_text(face = "bold", size = 6.5))
```

**Figure 1.** The 5 most abundant words in each chapter of "Capitalism VS Bolshevism". Word summary excluded stop words. 

```{r fig.height=6.5}
# bind text to nrc lexicon
text_nrc <- text_words_clean %>% 
  inner_join(get_sentiments("nrc"))

# find counts of words by sentiment bin
text_nrc_counts <- text_nrc %>% 
  count(chapter, sentiment) %>%
  # remove na values
  na.omit

# plot nrc sentiment counts 
ggplot(data = text_nrc_counts, 
       aes(x = sentiment, 
           y = n)) +
  # color columns
  geom_col(fill = "aquamarine4") +
  # facet wrap by chapter, free scales, relabel facet titles
facet_wrap(~chapter, scales = "free", labeller = labeller(chapter = chapters)) +
  # flip coordinates
  coord_flip() +
  # set theme
  theme_test() +
  # fix labs
  labs(y = " ", 
       x = " ", 
       title = "NRC Sentiment Occurence By Chapter") +
  # adjust text appearance 
  theme(plot.title = element_text(size = 18, hjust = 0.5), 
        strip.text.x = element_text(face = "bold", size = 6.5))
```

**Figure 2.** The relative abundance of 10 NRC sentiment associations to words in each chapter of "Capitalism VS Bolshevism". Word summary excluded stop words. 

### Analysis

- General patterns:
  - For the majority of chapters, "positive" and "trust" were the two most common sentiments, suggesting a generaly positive tone. 
  - The "negative" sentiment was still relatively abundant in most chapters, just to a lesser degree than "positive" and "trust" suggesting this work might switch back and forth between a "positive" and "negative" tone.  
  - The sentiments "fear" and "disgust" made up a larger proportion in the chapters "Bolshevism" and "Discontent" relative to other chapters, which calls for a more specific exploration of the Lexicon.  
  
- Potential Lexicon bias: 
  - Upon inspecting a few specific direct sentiment associations withing NRC Lexicon, some interesting patterns became apparent. 
  - The word "capitalism" was always associate with a "positive" sentiment. 
  - The words "socialism" and "socialist" were associated with the sentiments "fear", "disgust", "sadness", "negative", and "anger".
  - Though likely a powerful tool for many applications, this exploration of a political text using the NRC Lexicon revealed a clear political bias that strongly influences the sentiment analysis as a whole. 
  - Bias in this lexicon is possibly a result of it's crowd sourced origin, and might benefit from input provided by a more politically diverse group of people.

***

### Citations:

^1^ Walker, George Levi. Capitalism vs. Bolshevism. Boston, Mass., Dukelow & Walker Co., 1919. http://archive.org/details/capitalismvsbols00walk.

^2^ Crowdsourcing a Word-Emotion Association Lexicon, Saif Mohammad and Peter Turney, Computational Intelligence, 29 (3), 436-465, 2013.






